### 2019-10-24 manuscript/MS_trophToxNTO.tex
up change in files and models

\begin{center}
	\Large\textbf{
	Trophic transfer of anticoagulant rodenticides while managing rodent pests: the fine line between predator-prey regulation and pesticide-pest regulation
}\par
\end{center}

@@

\begin{abstract} ~
	\begin{enumerate}
		\item Understanding pesticide impacts on populations of target/non-target species and communities is a challenge to applied ecology. When predators that otherwise regulate pest densities ingest prey contaminated with pesticides, this can suppress predator populations as contaminated prey act as a super-predator, with pesticides controlling both pest and predators. It is, however, unknown how species relationships and protocols of rodenticide treatments, i.e. farmer functional responses, interact to affect pest regulation.
		\item Here we used linked differential equations to model a heuristic non-spatialized system including montane water voles, specialist vole predators (stoats, weasels), and a generalist predator (red fox) which consume voles, mustelids and other prey. By considering anticoagulant rodenticide toxicokinetic and toxicodynamic equations, we explored the impact of 5 farmer functional responses (defined by both rodenticide quantity and threshold vole density above which rodenticide spreading is prohibited) on predator-prey interactions, rodenticide transfer across the trophic chain and population effects.
		\item Spreading rodenticide while maintaining sufficient voles as prey resources led to less rodenticide being applied and extended periods without rodenticide in the environment, benefitting predators while avoiding episodes with high vole density. This may meet farm production interests while minimizing the impact on small mustelid and fox populations.
		Spreading rodenticide to maintain low vole densities suppressed mustelid and fox populations, leading to vole population dynamics being entirely regulated by rodenticide use. This vole-eradication treatment regime inhibited predator ecosystem services and promoted pesticide dependence.
		\item Farmer functional response with high or intermediate densities thresholds triggering treatment produced vole population fluctuations caused by rodenticide that could be followed by periods during which mustelids regulated vole populations. These alternative phases of mustelids and farmer regulation highlight the benefit of intraguild relationship where mustelids may rescue foxes from poisoning.
	\end{enumerate}

\paragraph{Synthesis and applications} Different farmer functional responses lead to a rich variety of population dynamics in predator-prey systems. That such a pesticide-tri-trophic system may cause a variety of population dynamics responses to pesticide use in agro-ecosystems is a novel insight. Our model reveals the need for maintaining refuges with sufficient non-poisoned voles for specialist mustelids, to conserve predator community, given the super-predator role of rodenticides. We suggest that long periods without pesticide treatment are essential to maintain predator populations, and that practices of pesticides use that attempt to permanently eradicate a pest over a large scale are counterproductive.

@@

Since the “green revolution” following the 1950s, pesticides use has increased to control pests damaging properties, public health or crops \citep{Tilman2002}. Pesticides usage is varyingly triggered by the perception/estimation of pest densities. Natural enemies (e.g., predators, parasites, competitors) also reduce pest densities and hence may preclude the need for using pesticides \citep{Michalko2017}. Natural enemies are, however, also affected by pesticides, either by direct exposure, through ingestion of contaminated prey \citep{Berny2007} or indirectly by cascade effects of resource depletion \citep{Halstead2014}.  Thus, under some regimes of pesticides use, pest populations only become regulated by pesticides once predators have collapsed. To preserve ecosystem health and the services of predators through regulation of pest densities, we need to assess the feasibility and benefit of pesticide treatment regimes in their ability to control pest species with minimal damage on predators \citep{Halstead2014}. It is however empirically challenging to assess the overall impact of pesticide treatments on the dynamics of species linked by trophic interactions. In this context, process-based models describing simplified scenarios are powerful tools to reveal hidden patterns by disentangling processes emerging from pesticide impacts on predator-prey system (e.g., \citet{Baudrot2018}).

@@

Voles and many small rodents are perceived as pests, but they are also keystone species, crucial to the functioning of grassland ecosystems, as well as being the prey of numerous predators, including species of conservation concern \citep{Delibes-Mateos2011, Coeurdassier2014}. Their population cycles create pulses of resources crucial to the viability of a wide range of resident predators and the aggregation of mobile avian vole predators \citep{Korpimaki1991}. The smallest mustelids (e.g. weasels Mustela nivalis) are specialist vole predators. Their numerical response has been shown theoretically to be necessary for generating predator-prey cycles \citep{Hanski1991}. They are said to be responsible for driving 3-to-5-year vole cycles in Fennoscandia (specialist predation hypothesis) \citet{Hanski1991}. Generalist resident predators like foxes (Vulpes vulpes) are expected to have regulatory and limiting effects on voles, owing to dietary plasticity that slows down vole population increase at low density \citet{Hanski1991}. Foxes do not show numerical responses to vole abundance \citep{Weber2002} but they influence the food chain through occasional killing and consumption of bite-sized mustelids. Mustelids form a small proportion of fox diet (0-10\%) but their offtake could represent a significant portion of the population (reviewed in \citet{Lambin2018}).

@@

With the aim of understanding the potentially complex interactions between prey that are perceived as pest predators and farmers spreading rodenticide in response to vole abundance and their functional responses, we studied a simplified system inspired by cyclically fluctuating montane water voles (\textit{Arvicola scherman}), small mustelids (stoats, weasels) that mostly eat voles (specialists), and foxes (generalists), with voles and mustelids as food items. We used a process-based model using differential equations to explore 5 FFR types of AR spread, combining population dynamics, predator-prey interactions and rodenticide transfer across the trophic chains. Model parameters and FFR were inspired by farming systems in the Jura Mountains, Franche-Comté (France), the region of Comté cheese production. In Franche-Comté, farmers shifted from polyculture to almost exclusively grass production for milk used to produce cheese from the early 1970s \citep{Giraudoux1997}. Due to recurrent vole outbreaks and damages to grasslands, massive rodenticide treatments were implemented from the early 80s with consequences on non-target wildlife. Practices developed technically under pressure from public opinion, farmer unions and farmer technical organizations collaborating with researchers to find treatment regimes with less harmful consequences for biodiversity \citep{Delattre2009}.

Hence, our main objective was to explore the properties of FFR in relation to varying vole density, to vole outbreak frequencies, to the guild of interacting predators and the effects on the global tri-trophic population dynamics. Additionally, we sought to establish under what regime of use the ARs deliver benefits to farmers without imposing damages to the ecosystem.

@@

The vole population followed a logistic growth rate, with $r_V$  the maximal reproduction rate, fixed at $r_V = \ln(2 \times 600)/365$ per day, since montane water vole populations can increase from 0  to  600 individuals ha$^{-1}$ or more \citep{Giraudoux1997} resulting in the equilibrium density being fixed at $K_V = 600$ individuals. 
%
The vole population was preyed upon by mustelid and fox populations. The vole consumption rate at different vole densities was described by functional responses ($\Phi_{V,M}$ for mustelids, $\Phi_{V,F}$  for foxes), see equations \eqref{eq:functionResponse}.
%
We treated small mustelids as vole specialist predators \citep{King2006}, assuming a Holling Type 2 functional response with attack rate $a_M$ and handling time $h_M$ (equation \eqref{eq:functionResponse}). We then represented foxes feeding on voles and mustelids by a multi-species functional response derived from Holling Type 3, referring to generalist feeding behaviour \citep{Baudrot2016}. For that, we denoted $a_{VF}$ and $a_{MF}$ the fox attack rate on voles and mustelids respectively. The parameter $h_F$ was the handling time for foxes.

@@

Figure~\ref{fig:scheme} represents the whole study system. Rodenticide is spread in grasslands during treatments, denoted $T_{Broma}(V)$, triggered by vole density $V$ (FFR described in Table~\ref{tab:results}).
%
Firstly, baits (50 mg kg$^{-1}$  of bromadiolone, hereafter AR) were spread in grasslands at quantity 7.5 to 20 kg ha$^{-1}$ day$^{-1}$. Notation day$^{-1}$ stands because of the daily time resolution.
%
Such quantity, $C$, was available for voles, and a proportion disappeared in the environment at rate $k_0$ (set at $k_0 = 0.0815$) \citep{Sage2008}. The proportion consumed per vole, with rate $\kappa(C)$, was assumed to be an increasing function. The function $\kappa(C)$ was characterized by a maximum ingestion $M_{in}$, and a half-saturation constant for ingestion $D_{in}$ in [mg kg$^{-1}$]:

@@

A fraction of AR ingested was accumulated in weasels while the rest was excreted with rate $k_{out,M}$.
%
AR provided to weasel though vole poisoning, is denoted $D_M$ induces lethal effect at rate $\mu_M(D_M)$, additive to natural mortality rate $m_M$. Parameter definition are further detailed in Supporting Information.
%
AR was ingested by foxes with a rate proportional to the functional response of foxes to voles, dead voles and mustelids.
%
Foxes also accumulated AR available in their prey, resulting in upward AR transfer in the trophic chain. Foxes accumulated AR in concentration $C_F$. A fraction of AR was excreted by foxes at rate $k_{out,F}$ at a rate between 0.38 and 0.72 [day$^{-1}$]\citep{Sage2010}, and AR caused fox mortality at rate $\mu_F$ ($D_F$).

We used log-logistic equations for describing dose-dependent mortality of animals exposed to AR. Vole and predator mortality rates due to AR $\mu_X(\zeta_X)$ ($X$ referring to the considered species, and $\zeta_X$ being the dose eaten $D_X$ or the internal concentration $D_X$) were expressed by equation (5):

\begin{equation}
\mu_X(\zeta_X) = \dfrac{1}{\text{experiment duration}} \times \left( 1- \dfrac{1}{1+ (LD_{50}/\zeta_X)^H} \right) 
\label{eq:muX}
\end{equation}

where $LD_{50}$ was the daily median lethal dose (50\% of population dying) with a value for voles of 2 mg kg$^{-1}$ as internal concentration, and a dose given as food of  2.1 mg kg$^{-1}$ mustelids \citep{Grolleau1989} and 0.5 mg kg$^{-1}$ for foxes \citep{Sage2010} (see Table~\ref{tab:parameterization}). Parameter $H$ is the Hill's coefficient, modulating the curve steepness and was estimated to fit the sparse data we have (Table~\ref{tab:parameterization}). Mortality rates considered the duration of AR toxicity from experiments of up to 6 days \citep{Sage2010}.

@@

We considered a range of realistic FFR spanning treatments during vole outbreaks only as a precautionary approach; in which treatments only takes place at intermediate or low vole density threshold. These scenarios are inspired by historic and contemporary protocols of bromadiolone use to control montane water voles in Franche-Comté, but also representative of practice globally (Table~\ref{tab:results}) \citep{Delattre2009}. To check the influence of foxes and intraguild predation on the system’s dynamics, we simulated scenarios with and without foxes.

Our simulations tracked the linked vole-mustelid-fox dynamics for 40 years, after a “burn-in” period of 10 years to reduce dependency of results upon initial conditions, to observe several vole cycles and to characterise AR effects on these species population dynamics. This burn-in period also had AR treatment triggered at specific vole densities and with a given rodenticide quantity for each FFR (see Table~\ref{tab:results}). The burn-in phase was selected according to a set of simulations with different initial conditions. Those simulations showed that in a given FFR (i.e., same threshold of vole density and amount of AR spread), the dynamics of the population were converging toward a similar pattern. 
In order to assess the impact of FFR on both agriculture and conservation interests, we estimated the following cost functions: (i) Number of treatment events per FFR ; (ii) Cumulative amount of AR (kg) ; (iii) Proportion of time when the AR-induced mortality of mustelids higher than 50\% (i.e. lethal exposure profile killing 50\% of mustelid population) ; (iv) Proportion of time when the mortality of mustelids was higher than 50\% due to natural mortality; (v) Proportion of time when the vole density was below 50 voles ha$^{-1}$, as a proxy for time when forage grass grows with low herbivore influence; (vi) Mean vole, mustelid and fox densities.

@@

All numerical analysis have been done using the Open language R and particularly the ODE solver package \textit{deSolve} \citep{Soetaert2010}. Model implementation and code to run analysis are available on the Github repository \url{https://github.com/virgile-baudrot/trophToxNTO/}.
%
From simulations, a sensitivity analysis allowed to study \textit{how uncertainties in the model output are apportioned to different sources of uncertainty in the model input} \citep{Saltelli2019}.
%
We applied a global approach using the first-order Sobol's sensitivity index $S_i$ defined as $S_i = \mathbb{V}(\mathbb{E}_{-i}(y \vert x_i)) / \mathbb{V}(y)$, where $\mathbb{V}(y)$ is the variance of $y$ when all factors are allowed to vary, $\mathbb{E}_{x\sim i}(y \vert x_i))$ the mean of $y$ when one factor is fixed \citep{Sobol1993}. We defined the domain of variation with a beta distribution within an hyperspace of plus or minus 10\% from the original value for all parameters.

@@

@@

Allowing for mortality by predators ingesting AR-poisoned voles changed the outcome of predator-prey dynamics involving vole, mustelid and fox populations. Secondary poisoning led to a rich spectrum of emergent dynamics according to the FFR to vole abundance.  
Without AR (FFR a), vole dynamics were regulated by mustelid predation that gave rise to 6-year cycles (Figure~\ref{fig:populations}). The fox population remained at its carrying capacity (i.e., 0.03 ind ha$^{-1}$, see Table~\ref{tab:parameterization}).

Under FFR b (high vole density threshold triggering treatment and high AR amount per treatment) and FFR d (intermediate threshold and low AR amount), vole dynamics were sequentially regulated by either AR treatments, which we refer to as farmer-regulated phase (red color sections in Figure~\ref{fig:populations}), or by mustelids, mustelids-regulated phase (blue color sections in Figure~\ref{fig:populations}). Farmer-regulated periods started when densities of living voles triggered treatments. This produced sudden declines of live voles followed by increases of dead voles. However, the vole population re-grew quickly which triggered frequent further treatments and pulses of availability of contaminated (both live and dead) voles. In FFR d (intermediate threshold and low AR), vole declines were not as deep as when pulses of AR amount were high (FFR b), owing to the reduced rodenticide amount per treatment. Mustelids and foxes also experienced AR-induced declines during this period. Under FFRs b (high threshold density, high AR) and d (intermediate threshold, low AR), mustelid-regulated periods started when mustelid numbers grew slowly to a peak, which depressed vole density, precluding rodenticide treatments and releasing the fox population from secondary poisoning, such that its abundance rebounded. Vole depletion by mustelids and subsequent mustelid declines allowed the vole population to grow again up to threshold densities and initiated a new period of regulation by farmers.
%
Vole dynamics were permanently regulated by AR treatment under FFR c (intermediate threshold, high AR, Figure~\ref{fig:populations}-c). Populations of live and dead voles experienced high frequency fluctuations (around 2 peaks every year) driven by AR. As AR treatments were frequent, being triggered by voles peaks, contaminated dead voles were always abundant (peaks at 90 ind ha$^{-1}$), and mustelid and fox densities remained low (see Supporting Information: mustelids <0.150 ind ha$^{-1}$ and foxes <0.01 ind ha$^{-1}$) but with a slowly increasing trend on the long term period (Figure~\ref{fig:populations}-c). This pattern of an increasing trend appears to be very sensitive to parameterization (i.e. a slight change in parameter may lead to a long-term decreasing pattern).
%
With FFR e (low threshold, low AR), vole populations were maintained by farmers at around 50 voles ha$^{-1}$ (See Figure~\ref{fig:populations} and Supporting Information). The population of live voles was buffered by treatments. Additionally, whenever voles reached densities triggering treatment, predator populations experienced strong declines. However, fox densities were higher compared to FFR c (intermediate threshold, high AR) (>0.01 ind ha$^{-1}$, see Supporting Information), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities. The maximum numbers of dead voles under this FFR e was relatively low (highest around 15 ind ha$^{-1}$) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles. This, in turn, induced mustelid and fox mortality and population declines (Supporting Information). Additionally, low availability of live voles triggered small mustelids mortality through starvation down to abundances similar to those resulting from AR use (Supporting Information).

\subsection{Comparison between farmer functional responses in terms of cost functions}

>>>>

%\begin{center}
%	\Large\textbf{
%	Trophic transfer of anticoagulant rodenticides while managing rodent pests: the fine line between predator-prey regulation and pesticide-pest regulation
%}\par
%\end{center}

\begin{center}
	\Large\textbf{
		Trophic transfer of pesticides: the fine line between predator-prey regulation and pesticide-pest regulation
	}\par
\end{center}

@@

\begin{abstract} ~
	\begin{enumerate}
		\item Understanding pesticide impacts on populations of target/non-target species and communities is a challenge to applied ecology. When predators that otherwise regulate pest densities ingest prey contaminated with pesticides, this can suppress predator populations as contaminated prey act as a super-predator, with pesticides controlling both pest and predators. It is, however, unknown how species relationships and protocols of anticoagulant rodenticide (AR) treatments, i.e. farmer functional responses, interact to affect pest regulation.
		\item Here we used linked differential equations to model a heuristic non-spatialized system including montane water voles, specialist vole predators (stoats, weasels), and a generalist predator (red fox) which consume voles, mustelids and other prey. By considering AR toxicokinetic and toxicodynamic, we explored the impact of 5 farmer functional responses (defined by both AR quantity and threshold vole density above which AR spreading is prohibited) on predator-prey interactions, AR transfer across the trophic chain and population effects.
		\item Keeping vole density below acceptable bounds by spreading AR while maintaining sufficient voles as prey resources led to less AR being applied and extended periods without AR in the environment, benefitting predators while avoiding episodes with high vole density. This may meet farm production interests while minimizing the impact on small mustelid and fox populations.
		Spreading AR to maintain low vole densities suppressed mustelid and fox populations, leading to vole population dynamics being entirely regulated by AR use. Such vole-eradication treatment regime inhibited predator ecosystem services and promoted pesticide dependence.
		\item An alternative population pattern emerges from a compromise between farmer functional response and population sensitivity to AR resulting in vole population fluctuations caused by AR and followed by periods during which mustelids regulated vole populations. These alternating phases of mustelids and farmer regulation highlight the benefit of intraguild relationship where mustelids may rescue foxes from poisoning.
		\item In such relatively simple mechanistic models, parameters encompass a large set of processes from the natural open systems used as reference. Both global and wide-range sensitivity analysis illustrated the delicacy to tune an optimal line between predator-prey regulation and pesticide-pest regulation.
	\end{enumerate}

\paragraph{Synthesis and applications} Different farmer functional responses lead to a rich variety of population dynamics in predator-prey systems. That such a pesticide-tri-trophic system may cause a variety of population dynamics responses to pesticide use in agro-ecosystems is a novel insight. Our model reveals the need for maintaining refuges with sufficient non-poisoned voles for specialist mustelids, to conserve the predator community, given the super-predator role of rodenticides. We suggest that long periods without pesticide treatment are essential to maintain predator populations, and that practices of pesticides use that attempt to permanently eradicate a pest over a large scale are counterproductive.

@@

Since the “green revolution” following the 1950s, pesticides use has increased to control pests damaging properties, public health or crops \citep{Tilman2002}. Pesticides usage is varyingly triggered by the perception/estimation of pest densities. Natural enemies (e.g., predators, parasites, competitors) also reduce pest densities and hence may preclude the need for using pesticides \citep{Michalko2017}. Natural enemies are, however, also affected by pesticides, either by direct exposure, through ingestion of contaminated prey \citep{Berny2007} or indirectly by cascade effects of resource depletion \citep{Halstead2014}.  Thus, under some regimes of pesticides use, pest populations only become regulated by pesticides once predators have collapsed. To preserve ecosystem health and the services of predators through regulation of pest densities, we need to assess the feasibility and benefit of pesticide treatment regimes in their ability to control pest species with minimal damage to predators \citep{Halstead2014}. It is however empirically challenging to assess the overall impact of pesticide treatments on the dynamics of species linked by trophic interactions. In this context, process-based models describing simplified scenarios are powerful tools to reveal hidden patterns by disentangling processes emerging from pesticide impacts on predator-prey system (e.g., \citet{Baudrot2018}).

@@

Voles and many small rodents are perceived as pests, but they are also keystone species, crucial to the functioning of grassland ecosystems, as well as being the prey of numerous predators, including species of conservation concern \citep{Delibes-Mateos2011, Coeurdassier2014}. Their population cycles create pulses of resources crucial to the viability of a wide range of resident predators and the aggregation of mobile avian vole predators \citep{Korpimaki1991}. The smallest mustelids (e.g. weasels \textit{Mustela nivalis}) are specialist vole predators. Their numerical response has been shown theoretically to be necessary for generating predator-prey cycles \citep{Hanski1991}. They are said to be responsible for driving 3-to-5-year vole cycles in Fennoscandia (the specialist predation hypothesis) \citet{Hanski1991}. Generalist resident predators like foxes (\textit{Vulpes vulpes}) are expected to have regulatory and limiting effects on voles, owing to dietary plasticity that slows down vole population increase at low density \citet{Hanski1991}. Foxes do not show numerical responses to vole abundance \citep{Weber2002} but they influence the food chain through occasional killing and consumption of bite-sized mustelids. Mustelids form a small proportion of fox diet (0-10\%) but their offtake could represent a significant portion of the population (reviewed in \citet{Lambin2018}).

@@

With the aim of understanding the potentially complex interactions between prey that are perceived as pest predators and farmers spreading rodenticide in response to vole abundance and their functional responses, we studied a simplified system inspired by cyclically fluctuating montane water voles (\textit{Arvicola scherman}), small mustelids (stoats, weasels) that mostly eat voles (specialists), and foxes (generalists), with voles and mustelids as food items. We used a process-based model using differential equations to explore 5 FFR types of AR spread, combining population dynamics, predator-prey interactions and rodenticide transfer across the trophic chain. Model parameters and FFR were inspired by farming systems in the Jura Mountains, Franche-Comté (France), the region of Comté cheese production. In Franche-Comté, farmers shifted from polyculture to almost exclusively grass production for milk used to produce cheese from the early 1970s \citep{Giraudoux1997}. Due to recurrent vole outbreaks and damages to grasslands, massive rodenticide treatments were implemented from the early 80s with consequences on non-target wildlife. Practices were developed technically under pressure from public opinion, farmer unions and farmer technical organizations collaborating with researchers to find treatment regimes with less harmful consequences for biodiversity \citep{Delattre2009}.

Hence, our main objective was to explore the properties of FFR in relation to varying vole density, to vole outbreak frequencies, to the guild of interacting predators and the effects on the global tri-trophic population dynamics. Additionally, we sought to establish what regime of use the ARs deliver benefits to farmers without imposing damages to the ecosystem.

@@

The vole population followed a logistic growth rate, with $r_V$  the maximal reproduction rate, fixed at $r_V = \ln(2 \times 600)/365$ per day, since montane water vole populations can increase from 0  to  600 individuals ha$^{-1}$ or more \citep{Giraudoux1997}, resulting in the equilibrium density being fixed at $K_V = 600$ individuals. 
%
The vole population was preyed upon by mustelid and fox populations. The vole consumption rate at different vole densities was described by functional responses ($\Phi_{V,M}$ for mustelids, $\Phi_{V,F}$  for foxes), see equations \eqref{eq:functionResponse}.
%
We treated small mustelids as vole specialist predators \citep{King2006}, assuming a Holling Type 2 functional response with attack rate $a_M$ and handling time $h_M$ (equation \eqref{eq:functionResponse}). We then represented foxes feeding on voles and mustelids by a multi-species functional response derived from Holling Type 3, referring to generalist feeding behaviour \citep{Baudrot2016}. For that, we denoted $a_{VF}$ and $a_{MF}$ the fox attack rate on voles and mustelids respectively. The parameter $h_F$ was the handling time for foxes.

@@

Figure~\ref{fig:scheme} represents the whole study system. Rodenticide is spread in grasslands during treatments, denoted $T_{Broma}(V)$, triggered by vole density $V$.
%
Firstly, baits (50 mg kg$^{-1}$  of bromadiolone, hereafter AR) were spread in grasslands at quantity 7.5 to 20 kg ha$^{-1}$ day$^{-1}$. Notation day$^{-1}$ stands because of the daily time resolution.
%
Such quantity, $C$, was available for voles, and a proportion disappeared in the environment at rate $k_0$ (set at $k_0 = 0.0815$) \citep{Sage2008}. The proportion consumed per vole, with rate $\kappa(C)$, was assumed to be an increasing function. The function $\kappa(C)$ was characterized by a maximum ingestion $M_{in}$, and a half-saturation constant for ingestion $D_{in}$ in [mg kg$^{-1}$]:

@@

A fraction of AR ingested was accumulated in weasels while the rest was excreted with rate $k_{out,M}$.
%
AR contaminating weasel though vole poisoning, is denoted $D_M$, induces lethal effect at rate $\mu_M(D_M)$, additive to natural mortality rate $m_M$. Parameter definitions are further detailed in Supporting Information.
%
AR was ingested by foxes with a rate proportional to the functional response of foxes to voles, dead voles and mustelids.
%
Foxes also accumulated AR available in their prey, resulting in upward AR transfer in the trophic chain. Foxes accumulated AR in concentration $C_F$. A fraction of AR was excreted by foxes at rate $k_{out,F}$ at a rate between 0.38 and 0.72 [day$^{-1}$]\citep{Sage2010}, and AR caused fox mortality at rate $\mu_F$ ($D_F$).

We used log-logistic equations for describing dose-dependent mortality of animals exposed to AR. Vole and predator mortality rates due to AR $\mu_X(D_X)$ ($X$ referring to the considered species) were expressed by equation (5):

\begin{equation}
\mu_X(\zeta_X) =   1- \dfrac{1}{1+ (LD_{50}/\zeta_X)^H} 
\label{eq:muX}
\end{equation}

where $LD_{50}$ was the daily median lethal dose (50\% of population dying), parameter $H$ is the Hill's coefficient modulating the curve steepness (see Table~\ref{tab:parameterization}).
%
Parameterization of $LD_{50}$ is important since it defines the inflexion point of the ecotoxicological effect. While it is a classical parameter targeted in experiments, its value can have different order of magnitude  \cite{Grolleau1989,Erickson2004,Sage2010}. We explored this parameterization through a wide-range sensitivity analysis detailed here-after (see also Figure\ref{fig:GSA}).

@@

We considered a range of realistic FFR spanning treatments during vole outbreaks only and a precautionary approach in which treatments only takes place at intermediate or low vole density threshold. These scenarios are inspired by historic and contemporary protocols of bromadiolone use to control montane water voles in Franche-Comté, but also representative of practice globally \citep{Delattre2009}: (A) a scenario without AR treatment, (B) a scenario with high vole density threshold triggering treatment (500 ind ha$^{-1}$) and high AR amount per treatment (20 kg ha$^{-1}$), (C) a third scenario with intermediate triggering threshold (250 ind ha$^{-1}$) and high AR per treatment (20 kg ha$^{-1}$), (D) scenario with intermediate threshold (250 ind ha$^{-1}$) and low AR amount (7.5 kg ha$^{-1}$) and (E) scenario with low threshold (50 ind ha$^{-1}$) and low treatment (7.5 kg ha$^{-1}$).
%
To check the influence of foxes and intraguild predation on the system dynamics, we also simulated scenarios with and without foxes.
%
Our simulations tracked the linked vole-mustelid-fox dynamics for 40 years, after a “burn-in” period of 10 years to reduce dependency of results upon initial conditions, to observe several vole cycles and to characterise AR effects on these species population dynamics. This burn-in period also had AR treatment triggered at specific vole densities and with a given rodenticide quantity for each FFR. The burn-in phase was selected according to a set of simulations with different initial conditions. Those simulations showed that in a given FFR (i.e., same threshold of vole density and amount of AR spread), the dynamics of the population were converging toward a similar pattern.

@@

All numerical analysis have been done using the open language R and particularly the ODE solver package \textit{deSolve} \citep{Soetaert2010}. Model implementation and code to run analysis are available on the Github repository \url{https://github.com/virgile-baudrot/trophToxNTO/}.

We applied a wide-range sensitivity analysis on lethal dose parameters (i.e. $LD_{50}$) for voles, mustelids and foxes, to explore changes in dynamical patterns.
%
While wide-range variations of each parameter is likely to have impact on the global pattern, the lethal dose parameter is a direct characteristic of the toxicological part of the model.
%
Since the $LD_{50}$ parameter of fox appears to have low impact on dynamical pattern in our model, we evaluated only 4 different values (i.e. 0.5, 1.5, 2 and 7 ppm) which are consistent with a large range covering such a parameter \cite{Erickson2004,Sage2010}.
%
For the prey and the specialist predator, this parameter was much more sensitive, so we explored 10 values for both, respectively in the range $[0.7-7]$ ppm and $[1-25]$ ppm following consistent ranges \cite{Grolleau1989,Erickson2004}. 
%
We classified the subsequent 1600 simulations according to the dynamical pattern of mustelids population: either "periodic regulation by mustelid" or "AR permanent regulation" where mustelids collapse.

Then, we performed a local sensitivity analysis to study changes in model outputs apportioned to small variation around model parameters input \citep{Saltelli2019}.
%
We applied a first-order Sobol's sensitivity index $S_i$ defined as $S_i = \mathbb{V}(\mathbb{E}_{-i}(y \vert x_i)) / \mathbb{V}(y)$, where $\mathbb{V}(y)$ is the variance of $y$ when all factors are allowed to vary, $\mathbb{E}_{x\sim i}(y \vert x_i))$ the mean of $y$ when one factor is fixed \citep{Sobol1993}. We defined the domain of variation with a beta distribution within an hyperspace of plus or minus 10\% from the original value for all parameters.
%
For output variable of local sensitivity analysis, we estimated the following cost functions: (i) Number of treatment events per FFR; (ii) Cumulative amount of AR (kg); (iii) Proportion of time when the AR-induced mortality of mustelids higher than 50\% (i.e. lethal exposure profile killing 50\% of mustelid population); (iv) Proportion of time when the mortality of mustelids was higher than 50\% due to natural mortality; (v) Proportion of time when the vole density was below 50 voles ha$^{-1}$, as a proxy for time when forage grass grows with low herbivore influence; (vi) Mean vole, mustelid and fox densities.

@@

Allowing for mortality by predators ingesting AR-poisoned voles changed the outcome of predator-prey dynamics involving vole, mustelid and fox populations. Secondary poisoning led to a rich spectrum of emergent dynamics according to the FFR to vole abundance.  
Without AR (scenario A), vole dynamics were regulated by mustelid predation that gave rise to a 6-year cycles (Figure~\ref{fig:populations}) typical of prey-specialist dynamics, while the generalist fox population remained at its carrying capacity (i.e., 0.03 ind ha$^{-1}$, see Table~\ref{tab:parameterization}).  
%
Based on this null model, we explored 4 scenarios of treatment described earlier and denoted respectively B, C, D and E, defined by the vole density threshold triggering treatment (respectively: 500, 250, 250 and 50 ind ha$^{-1}$) and the AR amount per treatment (respectively: 20, 20, 7.5 and 7.5 kg ha$^{-1}$).

\subsection{Sensitivity analysis of model parameters}

The exploration of ecotoxicological uncertainties on population dynamics has been done through the lethal dose parameter $LD_{50}$ for voles, mustelids and foxes (see Figure~\ref{fig:GSA}).
%
From the 1600 patterns explored in this wide-range sensitivity analysis, we observed two typical patterns. On one hand (i) vole dynamics were sequentially regulated by either AR treatments, which we refer to as farmer-regulated phase (white sections in Figure~\ref{fig:populations}), or by mustelids, mustelids-regulated phase (grey sections in Figure~\ref{fig:populations}). This pattern of succession of farmer-regulated and mustelids-regulated phases are denoted "Mustelid periodic regulation" in Figure~\ref{fig:GSA} (black pixels). On the other hand, (ii) vole dynamics were permanently regulated by AR treatments, because of crash of the mustelid population, what is referred as "Permanent AR regulation" in Figure~\ref{fig:GSA} (grey pixels).
%
We see that  $LD_{50}$ for foxes does not change the general pattern of dynamics.
Both scenario B and D exhibited the highest part of "Mustelid periodic regulation" compared to other scenarios C and E. This illustrated the fine compromise between resource and AR impact to benefit from predator regulation periods.

For other parameters, we performed a local sensitivity analysis where we computed a first order sensitivity index \citep{Sobol1993, Saltelli2019} given the contribution of each parameter to the variance (in Figure~\ref{fig:sobol_sensitivity} for scenario D - intermediate amount of 7.5kg AR at threshold density of 250 voles).
%
Parameters were moved within +/- 10\% interval around the fixed value as defined in scenario D (see Table~\ref{tab:parameterization}).
%
Crossing all cost functions to test the sensitivity of model parameters (Figure~\ref{fig:sobol_sensitivity}), the influence of each parameter appeared quite homogeneous.
%
There was no redundant or conversely any parameters with a dominant influence on model behaviour.
%
The homogeneity of the sensitivity to parameters was particularly true for densities of predators to which we focused our attention.

@@

While both "Mustelid periodic regulation" and "Permanent AR regulation" pattern  are possible in each of the 4 scenarios with AR (B, C, D and E) (see Figure~\ref{fig:GSA}), we selected for Figure~\ref{fig:populations} the most contrasting population dynamics for each of them in order to better describe their underlying mechanism. The contrasting border emerge for vole $LD_{50}$ around 1.5 to 2 ppm and in the range 2 to 2.25 ppm for mustelid $LD_{50}$.

From Figure~\ref{fig:populations}, we see that scenario B (high vole density threshold triggering treatment and high AR amount per treatment) and scenario D (intermediate threshold and low AR amount) present the most of dynamics with sequentially farmer-regulated phase and mustelids-regulated phase.
%
Farmer-regulated periods started when densities of living voles triggered treatments. This produced sudden declines of live voles followed by increases of dead voles. However, the vole population re-grew quickly which triggered frequent further treatments and pulses of availability of contaminated (both live and dead) voles.
%
Also, in scenario D, vole declines were not as deep as when pulses of AR amount were high like in scenario B (see curves in Supporting Information), owing to the reduced AR amount per treatment. 
%
Mustelids and sometimes foxes also experienced AR-induced declines during this period (Figure~\ref{fig:populations}-c,f,g,h and Figure~\ref{fig:intraguild}). For the predators, mustelid-regulated periods started when mustelid numbers grew slowly to a peak, which depressed vole density, precluding AR treatments and releasing the fox population from secondary poisoning, such that its abundance rebounded. Vole depletion by mustelids and subsequent mustelid declines allowed the vole population to grow again up to threshold densities and initiated a new period of regulation by farmers.
%
Sometimes, mustelids population has a slowly increasing trend (Figure~\ref{fig:populations}-g)). Under this regime, this increasing trend appears to be very sensitive to parameterization.

For scenario C (intermediate threshold and high AR) vole dynamics were mostly regulated by AR treatment only (e.g. Figure~\ref{fig:populations}-e). Populations of live and dead voles experienced high frequency fluctuations driven by AR. As AR treatments were frequent, being triggered by vole peaks, contaminated dead voles were always abundant (see Supporting Information: peaks at 90 ind ha$^{-1}$).
%
With scenario E (low threshold, low AR) vole populations were maintained by farmers at around 50 voles ha$^{-1}$ (See Figure~\ref{fig:populations} and Supporting Information).
%
The population of live voles was buffered by treatments. 
%
For most of simulations (except when vole $LD_{50}$ with high, Figure~\ref{fig:GSA}), whenever voles reached densities triggering treatment, predator populations experienced strong declines.
%
However, fox densities were higher compared to scenario C (intermediate threshold, high AR) (see Supporting Information), reflecting the reduced amount of AR used (7.5 kg) and transferred to foxes as there were lower vole densities.
%
The maximum numbers of dead voles under this scenario E was relatively low (highest around 15 ind ha$^{-1}$) but, due to frequent treatments, there was a steady replenishment of contaminated dead voles. This, in turn, induced mustelid and fox mortality and population declines (Supporting Information).
%
Additionally, low availability of live voles triggered small mustelids mortality through starvation, down to abundances similar to those resulting from AR use (Supporting Information).

%\subsection{Comparison between farmer functional responses in terms of cost functions}
